---
title: "figure_2AB"
author: "Cesar Arze (Ring Therapeutics)"
output:
 html_document:
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(ggplot2)
library(reshape2)
library(ggthemes)
library(ggpubr)
library(ggsci)
library(here)
library(magrittr)
library(patchwork)
```

## Fig 2: Characterization of the Personal Anellome (Panel A)
*Overview of study design. Fifteen transfusion recipients paired with one or more blood donors received a blood transfusion following surgery.*
*Recipient samples were collected post-transfusion over a period of 280 days. Pan-anellovirus PCR-positive samples in red.*


### Setup
***
Read in metadata file containing Pan-anello PCR status of all samples and counts of ORF1s and Anellovirus contigs per sample obtained from AnelloScope sequencing.

```{r, panel_a_setup}
timeline_metadata_df <- read.csv(here("data",
                                      "figure_1",
                                      "RingTx_transfusion1_metadata.csv"), header = TRUE)
timeline_metadata_df %<>% mutate(transfusion_set = gsub("^(R\\d{2})[T|D]\\d{2}.*", "\\1", RING_ID))
timeline_metadata_m <- timeline_metadata_df %>%
  filter(!(is.na(RING_ID))) %>% 
  distinct(RING_ID, day_of_draw, .keep_all = TRUE) %>%
  melt(id.vars = c("RING_ID", "transfusion_set"), measure.vars = c("day_of_draw")) %>%
  mutate(timepoint = case_when(grepl("D\\d{2}", RING_ID) ~ "donor",
                               TRUE ~ gsub("^R\\d{2}(T\\d{2})", "\\1", RING_ID)))

## Add some other metadata here some breakdowns about our patients (i.e. Sex or Age)
donor_recip_metadata_df <- read.csv(here("data",
                                         "figure_1",
                                         "RingTx_transfusion1_donor_recip_metadata.csv"), header = TRUE)
donor_recip_metadata_df %<>% mutate(category = case_when(grepl("R\\d{2}D\\d{2}", ID) ~ "donor",
                                                         TRUE ~ "recipient"))
donor_recip_metadata_df %<>% mutate(subject = gsub("^(R\\d{2})[T|D].*", "\\1", ID))
donor_recip_metadata_df %<>%   mutate(timepoint = case_when(grepl("D\\d{2}", ID) ~ "donor",
                               TRUE ~ gsub("^R\\d{2}(T\\d{2})", "\\1", ID)))

timeline_metadata_df %>% head()
donor_recip_metadata_df %>% head()
```

### Panel A

```{r, anello_pcr_status, echo=FALSE, fig.height=6, fig.width=18}

p1 <- donor_recip_metadata_df %>% 
  mutate(timepoint = gsub("^R\\d{2}([T|D]\\d{2}).*", "\\1", ID)) %>%
  filter(grepl("D\\d{2}", timepoint)) %>%
  mutate(subject = gsub("R(\\d{2})", "D\\1", subject)) %>%
  group_by(subject) %>%
  ggplot(aes(x = timepoint, y = subject, color = Anello.Status)) +
    geom_point(size = 5) +
    theme_fivethirtyeight() +
    ylab("Donors") + 
    scale_color_aaas(name = "Anellovirus PCR Status") +
    theme(plot.background = element_rect(fill = "white"),
          panel.background = element_rect(fill = "white"),
          strip.background.x = element_rect(fill = "white"),
          legend.background = element_rect(fill = "white"),
          legend.key = element_rect(fill = "white"),
          axis.text.x = element_blank(),
          axis.text.y = element_text(size = 16),
          axis.title.y = element_text(size = 18, face = "bold", vjust = 3),
          axis.title.x = element_blank(),
          legend.title = element_text(size = 15),
          legend.text = element_text(size = 15),
          legend.position = "none",
          plot.margin = unit(c(0.5, 0, 0, 0), "cm"))

p2 <- timeline_metadata_m %>%
  filter(timepoint != "donor") %>%
  left_join(donor_recip_metadata_df, by = c("RING_ID" = "ID")) %>%
  ggplot(aes(x = value, y = transfusion_set, color = Anello.Status)) +
    geom_point(size = 6) +
    xlab("Days Since Transfusion") +
    ylab("Recipients") +
    theme_fivethirtyeight() +
    scale_color_aaas(name = "Anellovirus PCR Status") +
    scale_x_continuous(breaks = seq(-20, 280, by = 20)) +
    theme(plot.background = element_rect(fill = "white"),
          panel.background = element_rect(fill = "white"),
          strip.background.x = element_rect(fill = "white"),
          legend.background = element_rect(fill = "white"),
          legend.key = element_rect(fill = "white"),
          axis.text.x = element_text(size = 16),
          axis.text.y = element_text(size = 16),
          axis.title = element_text(size = 18),
          axis.title.y = element_text(size = 18, face = "bold", vjust = 3),
          axis.title.x = element_text(vjust = -0.08),
          legend.title = element_text(size = 15),
          legend.text = element_text(size = 15),
          legend.position = "bottom")

g1 <- ggarrange(p1, p2, common.legend = T, legend = "bottom", align = "hv", ncol = 2, widths = c(.2, .8))
g1
```

### Panel B

``` {r, panel_B_setup, echo=FALSE}
transfusion_1_contig_orf1_stats <- read.csv(here("data",
                                                 "figure_1",
                                                 "RingTx_transfusion1_contig_and_orf_stats.csv"), header = TRUE)
transfusion_1_contig_orf1_stats %<>% mutate(transf_set = gsub("^(R\\d{2})[T|D].*", "\\1", RING_ID))
transfusion_1_contig_orf1_stats %<>% mutate(subject = case_when(grepl("R\\d{2}D\\d{2}", RING_ID) ~ RING_ID,
                                                                TRUE ~ gsub("^(R\\d{2})[T|D].*", "\\1", RING_ID)))
transfusion_1_contig_orf1_stats %<>% mutate(time_point = case_when(grepl("R\\d{2}D\\d{2}", RING_ID) ~ gsub("R\\d{2}(D\\d{2})", "\\1", RING_ID),
                                                                   TRUE ~ gsub("R\\d{2}(T\\d{2})", "\\1", RING_ID)))
transfusion_1_contig_orf1_stats %<>% mutate(category = case_when(grepl("R\\d{2}D\\d{2}", RING_ID) ~ "donor",
                                                                 TRUE ~ "recipient"))

transfusion_1_contig_orf1_stats %>% head()
```

```{r, donor_recipient_orf1_counts, echo=FALSE, fig.height=6, fig.width=18}

color_palette <- c('#a6cee3','#1f78b4','#b2df8a','#33a02c','#fb9a99',
                   '#e31a1c','#fdbf6f','#ff7f00','#cab2d6','#6a3d9a',
                   '#ffff99','#b15928','lightgrey','dimgrey')

recip_histo_p_low <- transfusion_1_contig_orf1_stats %>%
  filter(category == "recipient" & subject != "R01") %>%
  ggplot(aes(x = anello_orf1, y = subject, fill = subject, order = rev(time_point))) +
    geom_bar(stat = "identity", position = position_stack(reverse = TRUE)) +
    theme_fivethirtyeight() +
    scale_fill_manual(values = color_palette) +
    coord_cartesian(xlim = c(0,50)) +
    xlab("No. of Anello Strains") +
    ylab("Recipient") +
    theme(plot.background = element_rect(fill = "white"),
          panel.background = element_rect(fill = "white"),
          strip.background.x = element_rect(fill = "white"),
          legend.background = element_rect(fill = "white"),
          axis.text.x = element_text(size = 15),
          axis.text.y = element_text(size = 15),
          axis.title = element_text(size = 18, face = "bold"),
          axis.title.y = element_text(vjust = 1.5),
          axis.title.x = element_text(vjust = -0.05),
          strip.text = element_text(size = 15, face = "bold"),
          legend.position = "none",
          legend.direction = "vertical",
          legend.title = element_text(size = 15),
          legend.text= element_text(size = 13))

recip_histo_p_high <- transfusion_1_contig_orf1_stats %>%
  filter(category == "recipient" & subject != "R01") %>%
  ggplot(aes(x = anello_orf1, y = subject, fill = subject, order = rev(time_point))) +
  geom_bar(stat = "identity", position = position_stack(reverse = TRUE)) +
  theme_fivethirtyeight() +
  scale_fill_manual(values = color_palette) +
  coord_cartesian(xlim = c(130, 300)) +
  xlab("No. of Anello Strains") +
  ylab("Recipient") +
  theme(plot.background = element_rect(fill = "white"),
        panel.background = element_rect(fill = "white"),
        strip.background.x = element_rect(fill = "white"),
        legend.background = element_rect(fill = "white"),
        axis.text.x = element_text(size = 15),
        axis.text.y = element_blank(),
        axis.title = element_blank(),
        strip.text = element_text(size = 15, face = "bold"),
        legend.position = "none",
        legend.direction = "vertical",
        legend.title = element_text(size = 15),
        legend.text= element_text(size = 13),
        plot.margin = unit(c(0, 3, 0, -3), "mm"))

recip_orf_count_p <- ggarrange(recip_histo_p_low, recip_histo_p_high, align = "h", widths = c(.6, .4))

dcolor_palette <- c("#a6cee3", "#b2df8a", "#33a02c", "#fb9a99", "#e31a1c",
                    "#cab2d6", "#6a3d9a", "#b15928", 'lightgrey','dimgrey')

donor_histo_p_low <- transfusion_1_contig_orf1_stats %>%
  filter(category == "donor" & anello_orf1 > 0) %>%
  ggplot(aes(x = anello_orf1, y = subject, fill = transf_set)) +
  geom_bar(stat = "identity", position = position_stack(reverse = TRUE)) +
  theme_fivethirtyeight() +
  scale_fill_manual(name = "", values = dcolor_palette) +
  coord_cartesian(xlim = c(0,15)) +
  scale_x_continuous(breaks = seq(0, 15, by=5)) +
  xlab("No. of Anello Lineages") +
  ylab("Donors") +
  theme(plot.background = element_rect(fill = "white"),
        panel.background = element_rect(fill = "white"),
        strip.background.x = element_rect(fill = "white"),
        legend.background = element_rect(fill = "white"),
        axis.text.x = element_text(size = 15),
        axis.text.y = element_text(size = 15),
        axis.title = element_blank(),
        axis.title.y = element_text(size = 18, face = "bold", vjust = 1.5),
        strip.text = element_text(size = 15, face = "bold"),
        legend.position = "none",
        legend.title = element_text(size = 15),
        legend.text= element_text(size = 13),
        plot.margin = unit(c(0, 0, 0, 3), "mm"))

donor_histo_p_high <- transfusion_1_contig_orf1_stats %>%
  filter(category == "donor" & anello_orf1 > 0) %>%
  ggplot(aes(x = anello_orf1, y = subject, fill = transf_set)) +
  geom_bar(stat = "identity", position = position_stack(reverse = TRUE)) +
  theme_fivethirtyeight() +
  scale_fill_manual(name = "", values = dcolor_palette) +
  coord_cartesian(xlim = c(30,110)) + 
  scale_x_continuous(breaks = seq(30,110, by=30)) +
  theme(plot.background = element_rect(fill = "white"),
        panel.background = element_rect(fill = "white"),
        strip.background.x = element_rect(fill = "white"),
        legend.background = element_rect(fill = "white"),
        axis.text.x = element_text(size = 15),
        axis.text.y = element_blank(),
        axis.title = element_blank(),
        strip.text = element_text(size = 15, face = "bold"),
        legend.position = "none",
        legend.title = element_text(size = 15),
        legend.text= element_text(size = 13),
        plot.margin = unit(c(6, 0, 12.5, 3), "mm"))

donor_orf_count_p <- ggarrange(donor_histo_p_low, donor_histo_p_high, align = "h", widths = c(.7, .3))

g2 <- ggarrange(donor_orf_count_p, recip_orf_count_p, align = "h", widths = c(.3, .7))

layout <- "
####
AAAA
####
BBBB
"

g1 / g2 + plot_annotation(tag_levels = 'A') + plot_layout(design = layout) & theme(plot.tag = element_text(size = 36))
ggsave(here("figures",
            "figure_2AB.pdf"), units = "mm", height = 275, width = 650, device = cairo_pdf, scale = 1)
```