---
title: "figure_3A"
author: "Cesar Arze (Ring Therapeutics)"
output: html_document
---

```{r, setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(reshape2)
library(ggplot2)
library(ggthemes)
library(ggTimeSeries)
library(tidyverse)
library(scales)
library(RColorBrewer)
library(foreach)
library(here)
library(ggpubr)
library(cowplot)
library(here)
library(Cairo)
library(gridExtra)
library(magrittr)

parse_orf_abundances <- function(abundance_file, drop_zero=F, fill_missing=F) {
  orf_abundances.df <- read.csv(abundance_file, sep="\t", row.names = 1, header=TRUE)

  orf_abundances.df$orf <- rownames(orf_abundances.df)
  rownames(orf_abundances.df) <- NULL

  if (drop_zero == T) {
    orf_abundances.df <- orf_abundances.df %>% 
      select_if(negate(function(col) is.numeric(col) && sum(col) == 0))
  }

  orf_abundances.m <- orf_abundances.df %>%
    melt(id.vars=c("orf"))

  orf_abundances.m <- orf_abundances.m %>% 
    left_join(metadata_df, by = c("variable" = "RING_ID")) %>%
    mutate(timepoint_days = as.numeric(day_of_draw)) %>%
    select(orf, variable, timepoint_days, value) %>%
    mutate(timepoint = gsub("R\\d{2}([T|D]\\d{2}).*", "\\1", variable)) %>% 
    mutate(coverage = gsub(".*_cov_(\\d+.\\d+)_.*", "\\1", orf)) %>%
    mutate(source = ifelse(str_detect(orf, "^R\\d{2}D\\d{2}"), "donor", "recipient")) %>%
    mutate(sample = gsub("(R\\d{2})[D|T]\\d{2}[_|-].*", "\\1", orf))

  if (fill_missing == T) {
    if (nrow(orf_abundances.m) > 0) {
      orf_abundances.m <- fill_missing_timepoints(orf_abundances.m)
    } else {
      orf_abundances.m <- orf_abundances.m %>%
        ungroup() %>%
        mutate(source = as.character(source))
    }
  }

  return(orf_abundances.m)
}
```

## Multiple Anellovirus Lineages are Transmitted via Blood Transfusion (Panel A)
*Area graph showing relative abundance of anellovirus lineages longitudinally following blood transfusion.*
*Lineages colored in shades of red denote transmitted lineages from the donor(s) while shades of blue indicate resident anellovirus.*


### Setup

Read in metadata file containing translation of categorical time-points (T00, T01, T02, etc.) to actual days since transfusion. 

For each set of samples, parse the relative abundance files (in MPA format) to produce one large dataframe containing all
relative abundances for plotting. Additioanlly setup a list containing the number of donor lineagse for each relative abundance
dataframe to aid in coloring of steamgraphs.

```{r, panel_a_setup}

metadata_df <- read.csv(here("data", 
                             "figure_1", 
                             "RingTx_transfusion1_metadata.csv"), header = TRUE) %>%
  distinct(RING_ID, .keep_all = TRUE)

## We don't look at every single recipient since transmission of Anelloviruses was sparse in some of these
## subjects.
subjects = c("R04", "R05", "R06", "R07", "R08",
            "R09", "R11", "R12", "R13", "R14", "R15")
orf_category_counts <- data.frame(subject=character(), d_count=integer(), r_count=integer())

rel_abundances_dfs <- foreach(subject = subjects) %do% {
  rabundances_df <- parse_orf_abundances(here("data",
                                              "figure_3",
                                              paste(subject, ".rabund.final.txt", sep = "")), drop_zero = T)
  donor_count <- rabundances_df %>% 
      filter(!(grepl("R\\d{2}D\\d{2}", variable))) %>%
      filter(grepl("^R\\d{2}D\\d{2}", orf)) %>%
      distinct(orf, .keep_all = T) %>%
      nrow()
  recip_count <- rabundances_df %>% 
      filter(!(grepl("R\\d{2}D\\d{2}", variable))) %>%
      filter(grepl("^R\\d{2}T\\d{2}", orf)) %>%
      distinct(orf, .keep_all = T) %>%
      nrow()
  
  orf_category_counts %<>% add_row(subject = subject, d_count = donor_count, r_count = recip_count)

  return(rabundances_df)
}

```

### Panel A

```{r, fig.height=12, fig.width=17, panel_a, echo=FALSE}

subject_rabund_plots <- foreach(rel_abundance_df = rel_abundances_dfs) %do% {
  tgt_subject <- rel_abundance_df %>% head() %>% select(sample) %>% distinct() %>% pull()
  donor_count <- orf_category_counts %>% filter(subject == tgt_subject) %>% select(d_count) %>% pull()
  recip_count <- orf_category_counts %>% filter(subject == tgt_subject) %>% select(r_count) %>% pull()

  rel_abundance_df %>%
    filter(!(grepl("R\\d{2}D\\d{2}", variable))) %>%
    ggplot(aes(x = as.numeric(timepoint_days), y = value, fill = orf)) +
    geom_area(alpha=1, size=0.35, colour="black") +
    ggtitle(tgt_subject) +
    theme_fivethirtyeight() +
    scale_y_continuous(labels = scales::percent) +
    scale_fill_manual(values = c(colorRampPalette(brewer.pal(9, "Reds")[6:9])(donor_count),
                                 colorRampPalette(brewer.pal(8, "Blues")[3:8])(recip_count))) +
    theme(legend.position = "none",
          axis.text.y = element_text(size = 15),
          axis.text.x = element_text(size = 15),
          panel.background = element_rect(fill = "transparent"), # bg of the panel
          plot.background = element_rect(fill = "transparent", color = NA), # bg of the plot
          panel.grid.major = element_blank(), # get rid of major grid
          panel.grid.minor = element_blank(), # get rid of minor grid
          legend.background = element_rect(fill = "transparent"), # get rid of legend bg
          legend.box.background = element_rect(fill = "transparent")) # get rid of legend panel bg)
}

subject_rabund_figures <- do.call(grid.arrange, c(subject_rabund_plots, ncol=4))
cowplot::ggdraw(annotate_figure(subject_rabund_figures,
                                     left = text_grob("Relative Abundance\n(Percent of Anellovirus Reads)", size = 18, rot = 90, face = "bold"),
                                     bottom = text_grob("Days since Transfusion", size = 18, face = "bold"))) +
  theme_fivethirtyeight() +
  theme(plot.background = element_rect(fill = "white"),
        panel.background = element_rect(fill = "white"),
        strip.background.x = element_rect(fill = "white"),
        legend.background = element_rect(fill = "white"))
ggsave(here("figures",
            "figure_3.pdf"), units = "mm", height = 286, width = 615, device = cairo_pdf, scale = 1)
```

