---
title: "table_S4"
author: "Cesar Arze (Ring Therapeutics)"
output: html_document
---

```{r, setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(reshape2)
library(ggplot2)
library(ggthemes)
library(gghighlight)
library(tidyverse)
library(scales)
library(RColorBrewer)
library(foreach)
library(here)
library(ggpubr)
library(cowplot)
library(here)
library(yarrr)
library(Cairo)
library(gridExtra)
library(gt)
library(magrittr)
```

## Table S4: Blood Transfusion Transmitted Anello's 

The following analysis uses abundance data for donor anellovirus lineages to determine whether or not a novel lineage was transferred to a blood transfusion recipient. The input to this data is an abundance table in MPA format as shown below.

```{r, setup_input_parsing}
# Parse metadata file containing categorical timepoints (T01, T02) mapping to actual days since transfusion.
metadata_df <- read.csv(here("data",
                             "figure_1",
                             "RingTx_transfusion1_metadata.csv"), header = TRUE)

fill_missing_timepoints <- function(abundance_df) {
  ## Let's go ahead and add any missing timepoints into our dataframe so we can
  ## have a full time course here
  all_timepoints <- c("T00", "T01", "T02", "T03", "T04")
  
  incomplete_orfs <- abundance_df %>% 
    group_by(orf, sample) %>% 
    filter(value == 0 & grepl("T\\d{2}", timepoint)) %>%
    tally() %>% 
    filter(n < 5) %>% 
    distinct(orf)
  
  for (inc_orf in incomplete_orfs$orf) {
    sample <- gsub("(R\\d{2})[T|D]\\d{2}.*", "\\1", inc_orf)
    timepoint <- ifelse(grepl("D\\d{2}", inc_orf), "T00", gsub("R\\d{2}(T\\d{2}).*", "\\1", inc_orf))
    
    missing_timepoints <- abundance_df %>% 
      filter(orf == inc_orf) %>%
      .$timepoint %>%
      setdiff(all_timepoints, .)
    
    if (length(missing_timepoints) > 0) {
      missing_timepoints <- paste(sample, missing_timepoints, sep="")
      
      orf_vec <- rep(inc_orf, length(missing_timepoints))
      source_vec <- rep("missing", length(missing_timepoints))
      samples_vec <- rep(sample, length(missing_timepoints))
      timepoint_vec <- gsub("R\\d{2}", "", missing_timepoints)
      
      days_vec <- metadata_df %>%
        filter(RING_ID %in% missing_timepoints) %>%
        select(day_of_draw) %>%
        unique() %>%
        pull() %>%
        as.double()
      
      abundance_df %<>% 
        ungroup() %>%
        add_row(orf=orf_vec, sample=samples_vec, timepoint=timepoint_vec, timepoint_days=days_vec, source=source_vec)
    } 
  }
  
  return(abundance_df)
}

parse_orf_abundances <- function(abundance_file) {
  orf_abundances_df <- read.csv(abundance_file, sep="\t", row.names = 1, header=TRUE)

  orf_abundances_df$orf <- rownames(orf_abundances_df)
  rownames(orf_abundances_df) <- NULL
  
  orf_abundances_m <- orf_abundances_df %>%
    melt(id.vars=c("orf"))
    
  orf_abundances_m %<>%
    left_join(metadata_df, by = c("variable" = "RING_ID")) %>%
    mutate(timepoint_days = as.numeric(day_of_draw)) %>%
    select(orf, variable, timepoint_days, value) %>%
    mutate(timepoint = gsub("R\\d{2}([T|D]\\d{2}).*", "\\1", variable)) %>% 
    mutate(coverage = gsub(".*_cov_(\\d+.\\d+)_.*", "\\1", orf)) %>%
    mutate(source = ifelse(str_detect(orf, "^R\\d{2}D\\d{2}"), "donor", "recipient")) %>%
    mutate(sample = gsub("(R\\d{2})[D|T]\\d{2}[_|-].*", "\\1", orf))
  
  if (nrow(orf_abundances_m) > 0) {
    orf_abundances_m <- fill_missing_timepoints(orf_abundances_m)
  } else {
    orf_abundances_m <- orf_abundances_m %>%
      ungroup() %>%
      mutate(source = as.character(source))
  }
  
  return(orf_abundances_m)
}

# Parse a single file to demonstrate what our input looks like
R04_orf_abundances_df <- read.csv(here("data",
                                       "figure_3",
                                       "R04.rabund.final.txt"), sep="\t", header = TRUE)

# This is the melted data we process
R04_orf_abundances_df %>% head()
```

In order to get our transmitted and not-transmitted anellovirus lineages we need to parse each abundance file first, 
followed by identifying those anelloviruses that transmit and those that didn't.


```{r, get_transmitted_not_transmitted_anellos}

get_transmitted_orfs <- function(orf_abundances.m, 
                                 min_occur=3, 
                                 abund_thresh=.000000001) {
  orfs_matches_df <- orf_abundances_m %>% 
    filter(value >= abund_thresh) %>% 
    group_by(orf, timepoint) %>% 
    select(orf, timepoint, timepoint_days, value, sample, coverage) %>% 
    arrange(orf) %>% 
    distinct(orf, timepoint, .keep_all = TRUE)
  
  orfs_two_matches_df <- orfs_matches_df %>% 
    filter(grepl("T\\d{2}", timepoint)) %>%
    group_by(orf) %>% 
    tally() %>% 
    filter(n >= min_occur)
  orfs_multi_matches_df <- orfs_matches_df %>% 
    filter(orf %in% orfs_two_matches_df$orf)
  
  return(orfs_multi_matches_df)
}

get_non_transmitted_orfs <- function(abundance_df) {
  not_transmitted_df <- abundance_df %>%
    group_by(orf, sample) %>%
    filter(value == 0 & grepl("^R\\d{2}D", orf) & grepl("T\\d{2}", timepoint)) %>%
    tally() %>% 
    filter(n == 5) %>% 
    distinct(orf)
  
  return(not_transmitted_df)
}

transmitted_orfs_dfs <- list()
not_transmitted_orfs_dfs <- list()
ret <- foreach(sample = c("R02", "R03", "R04", "R05", "R06", "R07", "R08",
                          "R09", "R10", "R11", "R12", "R13", "R14", "R15")) %do% {
                            orf_abundances_m <- parse_orf_abundances(here("data",
                                                                          "figure_3",
                                                                          paste(sample, ".rabund.final.txt", sep = "")))
                            transmitted_orfs <- get_transmitted_orfs(orf_abundances_m, min_occur = 1, abund_thresh = 0.000001)
                            not_transmitted_orfs <- get_non_transmitted_orfs(orf_abundances_m)

                            transmitted_orfs_dfs <- c(transmitted_orfs_dfs, list(transmitted_orfs %>% ungroup()))
                            not_transmitted_orfs_dfs <- c(not_transmitted_orfs_dfs, list(not_transmitted_orfs))
}

transmitted_orf_anellos = bind_rows(transmitted_orfs_dfs)
not_transmitted_anellos = bind_rows(not_transmitted_orfs_dfs)

donor_distinct_timepoints <- transmitted_orf_anellos %>%filter(grepl("^R\\d{2}D", orf)) %>%
  select(timepoint) %>%
  distinct() %>%
  pull()

transmitted_orfs_df <- transmitted_orf_anellos %>%
  filter(grepl("^R\\d{2}D", orf)) %>%
  spread(key = timepoint, value = value) %>%
  group_by(orf) %>%
  fill(donor_distinct_timepoints, .direction = 'up')  %>% 
  slice(1) %>%
  select(-timepoint_days)
transmitted_orfs_df %>% head()

transmitted_orfs_df %>% 
  write.csv(here("data", "supplementary", "table_S4.csv"), row.names = F)

not_transmitted_anellos %>% head()
```

Now we can do some quick tallying of transmitted vs. not transmitted 

```{r, transmitted_stats}
transmitted_orf_anellos %>%
  filter(grepl("^R\\d{2}D", orf)) %>%
  group_by(sample) %>%
  distinct(orf) %>%
  tally()

transmitted_orf_anellos %>%
  filter(grepl("^R\\d{2}D", orf)) %>%
    group_by(orf) %>%
    tally() %>%
    arrange(-n)

## How many re-dosed Anello's did we have?
transmitted_orf_anellos %>%
  filter(grepl("^R\\d{2}D", orf)) %>%
  filter(timepoint == "T00" & value > 0)
```
